# Reproducible Research Course Project
This is R markdown document created for Peer Assignment project  related Reproducible Research Course

## Loading and processing the data
```{r}
library(dplyr)
library(lattice)
activity <- read.csv("activity.csv")

##creating new data set by removing NA records
activity_n <- filter(activity, !is.na(activity$steps)==TRUE) # removing step with NA records

##Converting the columns appropriately 
activity_n$date <- as.Date(as.character(activity_n$date))
activity_n$steps <- as.numeric(as.character(activity_n$steps))
```

##Calculating the Total Number of Steps taken per day?

```{r}
tot_steps <- aggregate(activity_n$steps, by=list(activity_n$date), FUN=sum)

##Giving proper names to data frame columns
names(tot_steps) <- c("Date", "Steps")

##Plotting a histogram
hist(tot_steps$Steps, col = "purple", breaks = 100,main = "Total Steps In a Day", xlab = "Steps")
rug(tot_steps$Steps)
```

##Calculating the Mean and Median of Steps taken per day
```{r}
##Calculating mean and median values and storing in new data frame
mn_mdn <- activity_n %>% group_by(date) %>% summarize(Mean = mean(steps), Median = median(steps))
```


##Plotting Time series plot  for Average 
```{r}
library(ggplot2)
##Creating new data frame 
interval <- aggregate(activity_n$steps, by= list(activity_n$interval), FUN=sum)
names(interval) <- c("Interval","Steps")
ggplot(interval, aes(Interval, Steps)) +geom_line(color ="blue") + xlab("5-minute Interval") + ylab("Average Number of Steps Taken")

```
##Calculating maximum of number of steps taken on 5-min interval
```{r}
interval[which.max(interval$Steps),]
```

##Missing Values
```{r}
##Missing steps
miss <- is.na(activity$steps)
table(miss)

##Filling missing value with mean step value for 5-min interval
activity_miss <- activity
activity_miss$steps[which(is.na(activity_miss$steps))] <- tapply(activity$steps, activity$interval, mean, na.rm = T, simplify = F)
activity_miss$steps <- as.vector(activity_miss$steps, mode = "numeric")

##Calculating total number of steps per day with new data set
tot_act_steps <- activity_miss %>% group_by(date) %>% summarize(total_steps= sum(steps))

##Histogram of the total number of steps taken each day
hist(tot_act_steps$total_steps, breaks = 10, col="pink", xlab = "Total Number of Steps", ylab = "Date", main = "Number of Steps taken Each Day")

##Calculating mean and median from the new data frame
act_mn_mdn <- activity_miss %>% group_by(date) %>% summarize(Mean = mean(steps), Median = median(steps))
```

##Difference Activity Pattern between Weekdays and Weekends

```{r}
## Converting the date column appropriately
activity_miss$date <- as.Date(as.character(activity_miss$date))

#Adding new column to identify Weekday and Weekend 
activity_miss$Day <- ifelse(weekdays(activity_miss$date) %in% c("Saturday","Sunday"),"Weekend","Weekday")

##Aggregating the data
agg_data <- aggregate(steps ~ interval + Day, activity_miss, mean)

##Plotting the activity pattern

ggplot(agg_data, aes(interval, steps)) + geom_line(color ="magenta") + facet_grid(Day ~.) + xlab("5-min interval") + ylab("Number of Steps")
```

